---
title: "A Priori Power Analysis for Visual Orientation Sensitivity"
author: "Rosewood & Seebold"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    theme: united
---

# Introduction

This document presents a simulated a priori power analysis for a **3x2x3 fully within-subjects repeated-measures design**. The analysis aims to determine the required sample size to detect a significant three-way interaction effect on two dependent variables: **Accuracy** and **Response Time (RT)**.

The experimental factors are:
1.  **Auditory Condition (3 levels):** Looming, Stationary, No Sound
2.  **Visual Field Location (2 levels):** Center, Periphery
3.  **Task Difficulty (3 levels):** Low, Medium, High

**Hypotheses:**
* **Center Looming Audio:** Enhanced visual orientation sensitivity for center audio compared to silent, indicated by higher accuracy and/or faster RT.
* **Side Looming Audio:** Asymmetric effects, with better performance when audio and visual stimuli are spatially congruent.
* **Peripheral Effect:** The facilitatory effect of auditory looming is more pronounced for peripheral visual targets.
* **Difficulty Dependence:** Impact of auditory looming varies with difficulty, with the most significant effects at intermediate levels.

# Method

The power analysis is conducted by simulating data based on effect sizes observed in a pilot study. For each sample size tested, data is simulated 1000 times, a repeated-measures ANOVA is performed on each simulated dataset, and the statistical power is calculated as the proportion of these simulations that yield a p-value below the alpha level (α = 0.05) for the three-way interaction.

## Setup

First, we load the necessary R packages for data manipulation (`tidyverse`), reading files (`readr`), and running ANOVAs (`afex`).

```{r setup, message=FALSE, warning=FALSE}
# Install packages if not already installed
if (!requireNamespace("tidyverse", quietly = TRUE)) install.packages("tidyverse")
if (!requireNamespace("afex", quietly = TRUE)) install.packages("afex")
if (!requireNamespace("readr", quietly = TRUE)) install.packages("readr")
if (!requireNamespace("readxl", quietly = TRUE)) install.packages("readxl")

library(tidyverse)
library(afex)
library(readr)
library(readxl)
```

## Step 1: Calculate Effect Sizes from Pilot Data

This chunk loads the pilot data from its full path, cleans it, and calculates the mean accuracy and response time for each of the 18 experimental conditions. These means (`mu_vector_acc` and `mu_vector_rt`) serve as the basis for the simulation.

```{r calculate_effect_sizes_from_pilot, message=FALSE, warning=FALSE}
# --- FIX: Provide the FULL path to the CSV file ---
# This ensures R can find the file regardless of the script's location.
# Please verify this is the correct path on your machine.
file_path <- "C:/Users/zach/OneDrive - North Carolina State University/TVLR/Looming_Sounds_Contrast/Zach work/ori_con_full.xlsx"
ori_con_data <- read_xlsx(file_path)

# Clean column names to be valid R variable names
names(ori_con_data) <- make.names(names(ori_con_data))

# Rename key columns, filter for the 'orientation' task, and create ordered factors.
# The key fix is renaming 'correct.' (created by make.names) to 'correct'.
analysis_data <- ori_con_data %>%
  dplyr::rename(
    difficulty = contrast.orientation,
    location = stimulus_direction,
    sound = looming_strength,
    rt = response_time,
    correct = correct # FIX: Standardize the 'correct' column name
  ) %>%  
  filter(task_type == "orientation") %>%
  mutate(
    sound_condition = factor(sound,
                             levels = c(32, 1, -1),
                             labels = c("Looming", "Stationary", "No_Sound")),
    visual_location = factor(location,
                             levels = c(0, 1),
                             labels = c("Center", "Periphery")),
    difficulty_level = factor(difficulty)
  )

# --- Calculate Mean Accuracy Vector ('mu_vector_acc') ---
accuracy_means <- analysis_data %>%
  group_by(sound_condition, visual_location, difficulty_level) %>%
  dplyr::summarise(mean_accuracy = mean(correct, na.rm = TRUE), .groups = 'drop') %>%
  arrange(sound_condition, visual_location, difficulty_level)
mu_vector_acc <- accuracy_means$mean_accuracy

# --- Calculate Mean Response Time Vector ('mu_vector_rt') ---
rt_means <- analysis_data %>%
  filter(correct == 1) %>% # Analyze RT only for correct responses
  group_by(sound_condition, visual_location, difficulty_level) %>%
  dplyr::summarise(mean_rt = mean(rt, na.rm = TRUE), .groups = 'drop') %>%
  arrange(sound_condition, visual_location, difficulty_level)
mu_vector_rt <- rt_means$mean_rt

# --- Display Vectors for verification ---
cat("--- For Accuracy ---\n")
cat("mu_vector_acc <- ")
dput(mu_vector_acc)

cat("\n--- For Response Time ---\n")
cat("mu_vector_rt <- ")
dput(mu_vector_rt)
```

## Step 2: Power Simulation for Accuracy

Here, we define the simulation parameters and run the power analysis for the **Accuracy** dependent variable. The data generation logic has been improved to correctly structure the within-subjects data.

```{r power_simulation_accuracy, cache=TRUE}
# Simulation Parameters
sample_sizes <- seq(30, 80, by = 5) # Test N from 30 to 80
nsim <- 1000 # Number of simulations per sample size
sd_acc <- 0.2 # Estimated standard deviation for accuracy
alpha <- 0.05 # Significance level

# Run the simulation loop
power_results_acc <- sapply(sample_sizes, function(n) {
  significant_results <- 0
  for (i in 1:nsim) {
    # FIX: Correctly structure the long-format data for repeated measures
    sim_data <- tibble(
      id = factor(rep(1:n, each = 18)),
      sound = factor(rep(c("Looming", "Stationary", "No_Sound"), each = 6, times = n)),
      location = factor(rep(c("Center", "Periphery"), each = 3, times = 3 * n)),
      difficulty = factor(rep(rep(1:3, each = 1), times = 6 * n)),
      accuracy = rnorm(n * 18, mean = rep(mu_vector_acc, times = n), sd = sd_acc)
    )
    
    # Run the ANOVA
    model_anova <- suppressMessages(aov_4(accuracy ~ sound * location * difficulty + (sound * location * difficulty | id), data = sim_data))
    
    # Check for significance of the 3-way interaction
    p_value <- summary(model_anova)$anova_table$`Pr(>F)`[7] # 7th row is the 3-way interaction
    if (!is.na(p_value) && p_value < alpha) {
      significant_results <- significant_results + 1
    }
  }
  significant_results / nsim
})

# Store results in a tibble
power_curve_data_acc <- tibble(
  n_subjects = sample_sizes,
  power = power_results_acc
)
```

# Results

### Accuracy Power Analysis Results

```{r acc_results_table, echo=FALSE, results='asis'}
if (exists("power_curve_data_acc")) {
  cat("\n#### Power Curve Results for Accuracy\n\n")
  print(knitr::kable(power_curve_data_acc, caption = "Simulated Power by Sample Size for Accuracy"))
  
  required_n_acc <- power_curve_data_acc %>%
    filter(power >= 0.80) %>%
    slice_min(n_subjects, n = 1) %>%
    pull(n_subjects)
  
  if (length(required_n_acc) > 0) {
    cat(paste0("\n**Required sample size (N) for 80% power (Accuracy) is approximately: ", required_n_acc, "**\n\n"))
  } else {
    cat("\n**Required sample size for 80% power (Accuracy) was not reached within the tested range.**\n")
  }
}
```

## Step 3: Power Simulation for Response Time (RT)

Next, we run the power analysis for the **Response Time** dependent variable, using the same corrected data structure.

```{r power_simulation_rt, cache=TRUE}
# Simulation Parameters
sd_rt <- 250 # Estimated standard deviation for RT in ms
# sample_sizes and nsim are reused from the accuracy simulation

# Run the simulation loop
power_results_rt <- sapply(sample_sizes, function(n) {
  significant_results <- 0
  for (i in 1:nsim) {
    # FIX: Correctly structure the long-format data for repeated measures
    sim_data <- tibble(
      id = factor(rep(1:n, each = 18)),
      sound = factor(rep(c("Looming", "Stationary", "No_Sound"), each = 6, times = n)),
      location = factor(rep(c("Center", "Periphery"), each = 3, times = 3 * n)),
      difficulty = factor(rep(rep(1:3, each = 1), times = 6 * n)),
      rt = rnorm(n * 18, mean = rep(mu_vector_rt, times = n), sd = sd_rt)
    )
    
    # Run the ANOVA
    model_anova <- suppressMessages(aov_4(rt ~ sound * location * difficulty + (sound * location * difficulty | id), data = sim_data))
    
    # Check for significance of the 3-way interaction
    p_value <- summary(model_anova)$anova_table$`Pr(>F)`[7]
    if (!is.na(p_value) && p_value < alpha) {
      significant_results <- significant_results + 1
    }
  }
  significant_results / nsim
})

# Store results in a tibble
power_curve_data_rt <- tibble(
  n_subjects = sample_sizes,
  power = power_results_rt
)
```

### Response Time (RT) Power Analysis Results

```{r rt_results_table, echo=FALSE, results='asis'}
if (exists("power_curve_data_rt")) {
  cat("\n#### Power Curve Results for Response Time\n\n")
  print(knitr::kable(power_curve_data_rt, caption = "Simulated Power by Sample Size for RT"))

  required_n_rt <- power_curve_data_rt %>%
    filter(power >= 0.80) %>%
    slice_min(n_subjects, n = 1) %>%
    pull(n_subjects)
  
  if (length(required_n_rt) > 0) {
    cat(paste0("\n**Required sample size (N) for 80% power (RT) is approximately: ", required_n_rt, "**\n\n"))
  } else {
    cat("\n**Required sample size for 80% power (RT) was not reached within the tested range.**\n")
  }
}
```

## Step 4: Plotting the Power Curves

Visualizing the results helps in understanding the relationship between sample size and statistical power.

```{r plot_power_curves, echo=FALSE, fig.cap="Power curves for Accuracy and Response Time"}
power_curve_data_acc$dv <- "Accuracy"
power_curve_data_rt$dv <- "Response Time"

combined_power_data <- bind_rows(power_curve_data_acc, power_curve_data_rt)

ggplot(combined_power_data, aes(x = n_subjects, y = power, color = dv)) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  geom_hline(yintercept = 0.8, linetype = "dashed", color = "red") +
  scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
  labs(
    title = "A Priori Power Analysis",
    subtitle = "Power to Detect a 3-Way Interaction Effect",
    x = "Sample Size (N)",
    y = "Statistical Power",
    color = "Dependent Variable"
  ) +
  theme_minimal()
```